Supporting Information for ``Robust Adaptation to Multi-Scale Climate Variability''

### 2019-04-04 report-earths-future/reviewer-response.md

# Response to Reviewers

We thank Dr. Montanari and one anonymous reviewer for their helpful critique of our paper "Robust Adaptation to Multi-Scale Climate Variability".

Major comments focused on three areas.
First, the key points were confusingly worded and uninformative.
Second, the theoretical discussion of low-frequency climate variability lacked connections to literature on the Hurst phenomenon.
Finally, more explicit links between the introduction, results, and discussion were needed.
We draw the attention of the reviewers to sections 1.2 and 2.4, which have been completely re-written to address these points.
We note that we have not applied track changes to these two sections, as they have been completely re-written.

We also thank the reviewers for thoughtful specific comments, discussed below.

@@

I would like to point out that I missed in the paper a discussion on the Hurst effect.
There is an extended literature on it. Such effect implies the presence of low frequency variability (long term cycles) which were observed in several secular time series of environmental variables.
The Hurst effect may be difficult to distinguish from anthropogenic climate change (ACC).
This effect is ignored by several climate studies. I believe this paper would be a good opportunity to discuss the interaction of the Hurst effect with ACC.
To this end, a proper discussion should be included in the manuscript.

I would like to congratulate the authors for their interesting work.

Best regards,
Alberto Montanari

> Thank you for your thoughtful and positive response. We share responses to your specific (line) comments below. We would like to point your attention to section 1.2, which has been substantially re-written and now provides a greater theoretical discussion of the Hurst phenomenon, low-frequency climate variability, fractals, and nonlinear dynamics.

@@

87: I fully agree with this sentence.

> Thank you very much for your important work on this topic

103: LFV is strictly related to the Hurst effect. There is an extended literature. I think a mention to the Hurst effect would be appropriate here.

> We have re-written this subsection to discuss literature on paleoclimate, Hurst effects, and ocean-atmosphere dynamics.

@@

118: This is why I believe we should not make reference to perfect models.

> Please see previous comment

231: I believe these details on the software should not be given here, in the body of the paper. These may be part of the user guide of the produced software

> We are deferring here to the editor. However, I feel that as a consumer of open source software I have an ethical responsibility to cite the primary packages that I use. Further, as a scientist committed to reproducibility I believe that describing software used is as important as describing the experiment design. However, it is true that making code accessible on GitHub partially alleviates this concern.

@@

## Reviewer 2

### Overall comments

The authors have teased apart climatic risks that are considered for planning purposes across multiple time horizons.
It is an interesting exploration the interaction of various types of climatic variability, planning horizons, and length of time series, and how this interaction impacts the choice of instruments of climate risk mitigation.
As this paper uses the word robust in the title, the introduction needs to link the concept of robustness to their questions and methods.
Suggest citing literature on robust adaptation if you keep ‘robust’ in the title.

> Thank you for this suggestion. Please see sentence beginning L45.

@@

15-16: Can you restate phrase copied below, unclear terms in bold:
**the role of low-frequency variability in modulating climate extremes**

> We agree and have re-phrased this sentence. It now reads "... historical and paleo climate records emphasize interannual to multidecadal modes of variability"

23: The discount rate is given as a decision factor but this is not modeled in the paper. I suggest relegating mention of a discount rate to the introduction or discussion.

> We have removed the distracting mention of a discount rate from this section.

@@

38-39: Is this formatting of reference correct?

> Per the document "Brief Guide to AGU Style and Grammar" document available at https://publications.agu.org/brief-guide-agu-style-grammar/, "AGU’s latex template and APA style in reference management software do not include this exception. You do not need to update your manuscript with this exception; it will be applied by the production vendor during the copy-editing process."

41: perhaps a good place to insert a definition and example of operational and financial instruments? An example is given for a structural instrument and that is what people are most likely to understand already.

> This is a helpful suggestion. The sentence now reads "Rather, actors such as New York city have turned to a combination of structural (e.g., a stormwater barrier), operational (e.g., improved evacuation routes), and financial (eg, index insurance) investments for reducing vulnerability and increasing resilience to climate extremes." We note that all of these examples are contained in the City of New York report referenced. We believe that these examples are sufficient since a strict definition of different classes of instruments would substantially disrupt the structure of the paragraph.

@@

53: comma after implications

> Added

67: perhaps a good place to define length of long-term short. Is this the interannual to multidecadal described in the abstract or the one-year term of a financial contract (line 59)?

> This is an important question, as we are certainly a bit vague here. The points are clearest in the limits of prediction at time scales less than one year or greater than a century; most real-world projects are in between. To add some clarity, we have added another sentence with an example. This sentence reads "For example, the action that New York City can take today which best protects against uncertain hurricane risks over the 21st century could potentially be to purchase insurance and defer more permanent allocation of capital until some uncertainties are resolved." We hope that our wording does not imply that we have performed any analysis for New York City.

@@

122: portend (singular)

> Fixed

Figure 1: Inconsistent capitalization. Can you add units?

> Thank you for this suggestion. Figure 1 now has units defined, uniform capitalization, and lettering (a)...(d).

@@

147: Recommend preserving the order you present the climate scenarios throughout the paper.

> Thank you for this helpful suggestion. We now order the scenarios (i) secular change only; (ii) LFV only; (iii) LFV plus secular change. This also mirrors the results.

### 2019-04-04 report-earths-future/paper-2018-robust-adaptation-cyclical-risk.tex
version for final coauthor comments

% Key Points
\begin{keypoints}
  \item \add{Climate risk varies over the planned lifespan of a climate risk mitigation instrument}
  \item 
\end{keypoints}

@@

Recent climate extremes such as floods, droughts, hurricanes, tornadoes, hailstorms, and heat waves have caused death and destruction, motivating investments in climate adaptation for the public and private sectors.
Further, rapid and continuing changes to global climate hazard and exposure underscore the need for adaptation strategies.
For example, population growth and urbanization have driven rapid increases in global exposure to events such as floods \citep{Jongman:2012cr} and tropical cyclones \citep{Peduzzi:2012iq}.
At the same time, anthropogenic modification of global and local climate processes affects the frequency, intensity, and location of extreme events \citep{IPCC:2012wt, Milly:2008dg, Shaw:2016bo}.
Even if \add{future} mitigation efforts are successful, \add{existing levels of atmospheric $\text{CO}_2$ and ocean heat content motivate the development of} novel adaptation strategies\remove{ are needed}.

This need has motivated a multitude of approaches for estimating the probability distribution of future climate risk, and for choosing between different risk mitigation instruments based on these estimates \citep[see, e.g.\@\xspace,][]{Merz:2014gf}.
\add{A typical goal is to create systems which are robust in the sense that they perform well over a wide range of plausible futures} \citep{lempert:2007,Borgomeo:2018hk} \add{and which fail along non-catastrophic modes} \citep{brown:2010}.
Although climate risk has been traditionally managed with centrally planned structural instruments (e.g.\@\xspace, a levee), the high price \citep{Papakonstantinou:2016ve}, environmental costs \citep{dugan:2010}, and vulnerability to biased climate projections \citep{lempert:2007} have recently slowed their implementation.
\remove{Instead}\add{Rather}, actors such as New York City have turned to a combination of structural\add{ (e.g.\@\xspace, stormwater barrier)}, operational\add{ (e.g.\@\xspace, improved evacuation routes)}, and financial\add{ (e.g.\@\xspace, index insurance)} instruments for reducing vulnerability and increasing resilience to climate extremes \citep{CityofNewYork:2013uh}.
These instruments are not typically implemented in isolation or statically.
Instead, investment decisions made at each point in time affect the viability, costs, and benefits of future decisions, causing the system to trace a ``pathway'' through time \citep{Walker:2013gi,Haasnoot:2013im,Haasnoot:2015dh}.

@@

Secular change is not the only mechanism which can cause historical records to provide a biased view of future risk.
The Hurst phenomenon associated with stochastic long-memory processes and has been invoked to explain long scale fluctuations in geophysics, physics, biology, medicine, traffic, network dynamics as well as for financial time series.
This phenomena can be characterized as:
\begin{equation}\label{eq:hurst}
  \frac{R_n}{S_n} \sim n^H,
\end{equation}
where $H \in [0.5, 1]$ is the Hurst coefficient, $R_n$ is the range of the first $n$ cumulative deviations from the mean, and $S_n$ is the sample standard deviation \citep{Hurst:1951tc,oconnell:2016,klemes:1974}.
\Cref{eq:hurst} measures the tendency of a time series to either regress to the mean or cluster towards the extremes.
The Hurst exponent $H$ is directly related to the fractal dimension of a process, and there is a rich multi-disciplinary literature as to the process level and statistical justification of long memory and fractal processes, especially in hydrology \citep{mandelbrot:1985,mandelbrot:1969,beran:1994}.
\citet{lovejoy:2012,lovejoy:2013a} and \citet{lovejoy:2013} also argue that multifractal and long memory processes are an appropriate description of climate in general and weather extremes in particular. 
Based on this theory, regimes and quasi-periodic phenomena with amplitude that increases with recurrence interval would be the norm rather than the exception.
Similar arguments are made by \citet{koutsoyiannis:2003}, \citet{markonis:2013}, and \citet{oconnell:2016}, and detailed discussions and some philosophical perspectives are offered by \citet{bloschl:2010} and \citet{montanari:2003}.
Wavelet methods have also been used to estimate the Hurst exponent \citep{simonsen:1998,chamoli:2007}, and to design simulation algorithms that reproduce self-similarity, long range dependence and quasi-periodic regimes \citep{Kwon:2007fj,bullmore:2001,geweke:1983,feng:2005}.

Links between local hydrological processes and large-scale climate patterns, often referred to as teleconnections \citep{Angstrom:1935ej}, have also been used to explain \gls{lfv} in hydroclimate systems.
If these climate modes vary on decadal or longer time scales, such as the \gls{pdo} \citep{Mantua:1997kj}, then this \gls{lfv} may appear in local hydrologgical records as well, leading the past to be a poor predictor of future behavior.
For example, \citet{Jain:2001hz} demonstrated that in a system with substantial \gls{lfv}, due to \gls{enso}, the past $N$ years can be a poor proxy for the following $N$ years even when no secular change is present.
When records are short and low-frequency external forcing or local memory dominates, it is easy to mistake the Hurst phenomenon for a secular trend which may not continue from the past into the future \citep{bhattacharya:1983}.
t is also difficult to identify the dominant scales of variability \citep{newman:2016,Jin:1994wq}.

We illustrate this type of \gls{lfv} in \cref{fig:observed-lfv}, which shows a \SI{500}{year} drought reconstruction from the \gls{lbda} \citep{Cook:2010bz}, a \SI{100}{year} record of annual maximum streamflowflow on the American River at Folsom, and the global wavelet power spectrum for both \citep{Torrence:1998jp,Roesch:wlBQQoIs}.
Peaks for the American River time series are apparent at approximately 2 and 15 years and in the \gls{lbda} time series at approximately 8, 20, and 64 years.
This is illustrated by the blue line in \remove{the top left panel of} \cref{fig:observed-lfv}\add{(b)}, which shows a \SI{20}{year} moving average of the \gls{lbda} time series.
A detailed analysis of these time series is beyond the scope of this paper, but we note that the high amplitude and long time periods of the quasi-periodic oscillations they exhibit are consistent with analyses of \gls{lfv} in other hydroclimate systems \citep{Kiem:2002kq,Swierczynski:2012km,Woollings:2014kd,Hodgkins:2017hw}.

@@

\Cref{fig:flowchart} describes the experimental design.
We assess estimation bias and variance for three \remove{cases}\add{scenarios of future climate}.
First, we consider an idealized scenario where only secular change is present in the system and \gls{lfv} is fully damped.
Next, we consider the ``pre-industrial'' case where there is no secular change but \gls{lfv} modulates climate risk in time.
Finally, we consider a more realistic (though still idealized) case with both \gls{lfv} and secular change.
Model parameters for each scenario are given in the supplemental methods section of the online supporting materials.

@@

Computation was carried out in the python programming language, making particular use of the matplotlib, numpy, pandas, pomegranate, scipy, and xarray libraries for scientific computing \citep{Hunter:2007ih, vanderWalt:2011dp, McKinney:2010un, schreiber:2017, Jones:2001uv, Hoyer:2017hs}.
Wavelet analysis was conducted using the WaveletComp package \citep{Roesch:wlBQQoIs} in the R programming language.
Bayesian models were written in the stan probabilistic programming language \citep{Carpenter:2017ke} using the No U-Turn Sampler \citep{Hoffman:2011wm,Betancourt:2017vd}.
The codes used to generate the figures and text of this paper are available at \url{https://github.com/jdossgollin/robust-adaptation-cyclical-risk} or in the online supplemental material of this paper.\footnote{Dear Reviewers: our codes are not presently available on GitHub, as we have chosen to keep the repository private until the article is accepted for publication, but they are available to you through the online supplemental material for this submission.}

@@

\Cref{fig:lfv-secular-nino3-bias-variance,fig:lfv-secular-nino3-bias-variance,fig:secular-nino3-bias-variance} \add{show that} for projects where $M$ is sufficiently short, intrinsic uncertainty is low and cyclical climate variability is dominant over the project planning period \citep{Jain:2001hz,Hodgkins:2017hw}.
However, one's \emph{ability} to identify and predict this variability depends on having a model of sufficient complexity to represent the processes that cause \gls{lfv}, and the data to fit the model.
In this case, the project may be in the ``potential predictability zone'' of \cref{fig:m-n-sketch}.
If sufficient information is not available, however, then simple models which represent fewer processes may be preferred (the ``rough guess zone'').

>>>>

% Key Points
\begin{keypoints}
  \item \add{Climate risk varies over the planned lifespan of a climate risk mitigation instrument}
  \item \add{We carry out stylized computational experiments to assess the bias and variance of future risk estimated from limited data sets}
  \item \add{Lack of data limits the identifiability of climate signals, favoring prediction of short-term risks which require less extrapolation}
\end{keypoints}

@@

Recent climate extremes such as floods, droughts, hurricanes, tornadoes, hailstorms, and heat waves have caused death and destruction, motivating investments in climate adaptation for the public and private sectors.
Further, rapid and continuing changes to global climate hazard and exposure underscore the need for adaptation strategies.
For example, population growth and urbanization have driven rapid increases in global exposure to events such as floods \citep{Jongman:2012cr} and tropical cyclones \citep{Peduzzi:2012iq}.
At the same time, anthropogenic modification of global and local climate processes affects the frequency, intensity, and location of extreme events \citep{IPCC:2012wt, Milly:2008dg, Shaw:2016bo}.
Even if \add{future} mitigation efforts are successful, \add{existing levels of atmospheric $\text{CO}_2$ and ocean heat content necessitate the development of} novel adaptation strategies\remove{ are needed}.

This need has motivated a multitude of approaches for estimating the probability distribution of future climate risk, and for choosing between different risk mitigation instruments based on these estimates \citep[see, e.g.\@\xspace,][]{Merz:2014gf}.
\add{A typical goal is to create systems which are robust in the sense that they perform well over a wide range of plausible futures} \citep{lempert:2007,Borgomeo:2018hk} \add{and which fail along non-catastrophic modes} \citep{brown:2010}.
Although climate risk has traditionally been managed with centrally planned structural instruments (e.g.\@\xspace, a levee), the high price \citep{Papakonstantinou:2016ve}, environmental costs \citep{dugan:2010}, and vulnerability to biased climate projections \citep{lempert:2007} have recently \remove{slowed their implementation}\add{dampened enthusiasm}.
\remove{Instead}\add{Rather}, actors such as New York City have turned to a combination of structural\add{ (e.g.\@\xspace, stormwater barrier)}, operational\add{ (e.g.\@\xspace, improved evacuation routes)}, and financial\add{ (e.g.\@\xspace, index insurance)} instruments for reducing vulnerability and increasing resilience to climate extremes \citep{CityofNewYork:2013uh}.
These instruments are not typically implemented in isolation or statically.
Instead, investment decisions made at each point in time affect the viability, costs, and benefits of future decisions, causing the system to trace a ``pathway'' through time \citep{Walker:2013gi,Haasnoot:2013im,Haasnoot:2015dh}.

@@

Secular change is not the only mechanism which can cause historical records to provide a biased view of future risk.
The Hurst phenomenon is very well known and identified with long memory processes in geophysics, physics, biology, medicine, traffic, network dynamics, and finance \citep{oconnell:2016}.
The extensive observations of such behavior in hydrologic and climatic time series emphasize the need to consider such processes as underlying any discussion of climate change or nonstationarity \citep{koutsoyiannis:2003,markonis:2013,Palmer:1993ef}.
The Hurst phenomena has also been connected to low frequency quasi-periodic phenomena, especially where fractal scaling is expected.
For example, wavelet methods have been used to estimate the Hurst exponent \citep{simonsen:1998,chamoli:2007}, and to design simulation algorithms that reproduce self-similarity, long range dependence and quasi-periodic regimes \citep{Kwon:2007fj,bullmore:2001,geweke:1983,feng:2005}.
The Hurst phenomenon also provides a link between catchment hydrology and global climate dynamics \citep{bloschl:2010,montanari:2003}.
The Hurst exponent is directly related to the fractal dimension of a process, and there is a rich multi-disciplinary literature as to the process level and statistical justification of long memory and fractal processes in hydrology \citep{mandelbrot:1985,mandelbrot:1969,beran:1994}.
These processes have also been used to describe multi-scale dynamics of the climate \citep{lovejoy:2012,lovejoy:2013a,lovejoy:2013,selvam:2017}, including \gls{enso} \citep{maruyama:2018,zivkovic:2013}.

External forcing from structured climate signals \citep[``teleconnections'';][]{Angstrom:1935ej} and catchment dynamics are both useful in explaining the \gls{lfv} observed in natural hydroclimate time series.
We illustrate such \gls{lfv} in \cref{fig:observed-lfv}, which shows a \SI{500}{year} drought reconstruction from the \gls{lbda} \citep{Cook:2010bz}, a \SI{100}{year} record of annual maximum streamflowflow on the American River at Folsom, and the global wavelet power spectrum for both \citep{Torrence:1998jp,Roesch:wlBQQoIs}.
Peaks for the American River time series are apparent at approximately 2 and 15 years and in the \gls{lbda} time series at approximately 8, 20, and 64 years.
This is illustrated by the blue line in \remove{the top left panel of} \cref{fig:observed-lfv}\add{(b)}, which shows a \SI{20}{year} moving average of the \gls{lbda} time series.
A detailed analysis of these time series is beyond the scope of this paper, but we note that the high amplitude and long time periods of the quasi-periodic oscillations they exhibit are consistent with analyses of \gls{lfv} in other hydroclimate systems \citep{Kiem:2002kq,Swierczynski:2012km,Woollings:2014kd,Hodgkins:2017hw}.
The key implication is that the observations, \citep{Jain:2001hz}, trends \citep{bhattacharya:1983}, and frequencies \citep{newman:2016} observed in the past are often poor predictors of future behavior.

@@

\Cref{fig:flowchart} describes the experimental design.
We assess estimation bias and variance for three \remove{cases}\add{scenarios of future climate}.
First, we consider an idealized scenario where only secular change is present in the system and \gls{lfv} is fully damped (``secular change only'').
Next, we consider the ``pre-industrial'' case where there is no secular change but \gls{lfv} modulates climate risk in time (``\acrlong{lfv} only'').
Finally, we consider a more realistic (though still idealized) case with both \gls{lfv} and secular change (``\acrlong{lfv} plus secular change'').
Model parameters for each scenario are given in the supplemental methods section of the online supporting materials.

@@

Computation was carried out in the python programming language, making particular use of the matplotlib, numpy, pandas, pomegranate, scipy, and xarray libraries for scientific computing \citep{Hunter:2007ih, vanderWalt:2011dp, McKinney:2010un, schreiber:2017, Jones:2001uv, Hoyer:2017hs}.
Wavelet analysis was conducted using the WaveletComp package \citep{Roesch:wlBQQoIs} in the R programming language.
Bayesian models were written in the stan probabilistic programming language \citep{Carpenter:2017ke} using the No U-Turn Sampler \citep{Hoffman:2011wm,Betancourt:2017vd}.
The codes used to generate the figures and text of this paper are available at \url{https://github.com/jdossgollin/2018-robust-adaptation-cyclical-risk} or in the online supplemental material of this paper.

@@

\Cref{fig:lfv-secular-nino3-bias-variance,fig:lfv-secular-nino3-bias-variance,fig:secular-nino3-bias-variance} \add{show that} for projects where $M$ is sufficiently short, intrinsic uncertainty is low and cyclical climate variability is dominant over the project planning period \citep{Jain:2001hz,Hodgkins:2017hw}.
However, one's ability to identify and predict this variability depends on having a model of sufficient complexity to represent the processes that cause \gls{lfv}, and the data to fit the model.
In this case, the project may be in the ``potential predictability zone'' of \cref{fig:m-n-sketch}.
If sufficient information is not available, however, then simple models which represent fewer processes may be preferred (the ``rough guess zone'').

### 2019-04-04 report-earths-future/si_template_2019.tex
version for final coauthor comments

As described in Section 2, we implement the stationary and trend log-normal models in a Bayesian framework using the stan programming language \cite{Carpenter:2017ke}.
The full model, including priors, for the stationary log-normal model is given by equation \eqref{log-normal-stationary}:
\begin{align}\label{eq:log-normal-stationary}
  \begin{split}
    \log Q_\text{hist} & \sim \mathcal{N} \qty(\mu, \ \sigma) \\
    \mu &\sim \mathcal{N} \qty(7, 1.5) \\
    \sigma &\sim \mathcal{N}^+ \qty(1, 1)
  \end{split}
\end{align}
where $\mathcal{N}$ denotes the normal distribution and $\mathcal{N}^+$ denotes a half-normal distribution.
The full model, including priors, for the trend log-normal model is given by equation \eqref{eq:log-normal-trend}:
\begin{align}\label{eq:log-normal-trend}
  \begin{split}
    \mu &= \mu_0 + \beta_\mu \qty(t - t_0) \\
    \log Q_\text{hist} & \sim \mathcal{N} \qty(\mu, \ \xi \mu) \\
    \mu_0 & \sim \mathcal{N} \qty(7, 1.5) \\
    \beta_\mu & \sim \mathcal{N} \qty(0, 0.1) \\
    \log \xi & \sim \mathcal{N} \qty(0.1, 0.1)
  \end{split}
\end{align}
where $\xi$ is an estimated coefficient of variation.

@@

In this section we describe the parameters used to generate the specific sets of sequences that are analyzed in the results section.
To explore the temporal structure of risk, we generate $J=1000$ sequences of streamflow, for each of several combinations of $M$ and $N$.
Streamflow sequences are generated using the stationary generating functions both with and without superimposing a secular trend.

@@

In the main text we sampled climate risk by considering dependence on a single climate index, as shown in equation 5.
Here we examine the sensitivity of our results to this assumption by considering a different parameterization for $\mu(t)$, specifically a two-state Markov chain model.
A Markov chain explicitly models transition between a fixed number of regimes, mimicking similar phenomena observed in nature.
The transition matrix is given by
\begin{equation}
  P = \mqty[\pi_1 & 1-\pi_1 \\ 1-\pi_2 & \pi_2].
\end{equation}
The transition matrix is first used to generate a sequence of states $S(t)$.
The value $\mu(t)$ depends only on $S(t)$ and on time itself:
\begin{equation}
  \mu(t) = \begin{cases}
    \mu_{1} + \gamma_1 \qty(t - t_0) & \qqtext{if} S(t) = 1 \\
    \mu_{2} + \gamma_2 \qty(t - t_0) & \qqtext{if} S(t) = 2
  \end{cases}
\end{equation}
For simplicity, we assume that the coefficient of variation is the same for both states and that $\pi_1=\pi_2$.
We further impose $\mu_{1} > \mu_{2}$ so that state 1 can be interpreted as the ``wet'' state and state 2 as the ``dry'' state.
Parameters for the three experiments are given in Table \ref{tab:markov-stationary}.
Results for these experiments are shown in Figure \ref{fig:lfv-markov-bias-variance}, \ref{fig:lfv-secular-markov-bias-variance}, and \ref{fig:secular-only-markov-bias-variance}.

>>>>

As described in Section 2, we implement the stationary and trend log-normal models in a Bayesian framework using the stan programming language \cite{Carpenter:2017ke}.
The full model, including priors, for the stationary log-normal model is given by equation \eqref{eq:log-normal-stationary}:
\begin{align}\label{eq:log-normal-stationary}
  \begin{split}
    \log Q_\text{hist} & \sim \mathcal{N} \qty(\mu, \ \sigma) \\
    \mu &\sim \mathcal{N} \qty(7, 1.5) \\
    \sigma &\sim \mathcal{N}^+ \qty(1, 1)
  \end{split}
\end{align}
where $\mathcal{N}$ denotes the normal distribution and $\mathcal{N}^+$ denotes a half-normal distribution.
The full model, including priors, for the trend log-normal model is given by equation \eqref{eq:log-normal-trend}:
\begin{align}\label{eq:log-normal-trend}
  \begin{split}
    \mu &= \mu_0 + \beta_\mu \qty(t - t_0) \\
    \log Q_\text{hist} & \sim \mathcal{N} \qty(\mu, \ \xi \mu) \\
    \mu_0 & \sim \mathcal{N} \qty(7, 1.5) \\
    \beta_\mu & \sim \mathcal{N} \qty(0, 0.1) \\
    \log \xi & \sim \mathcal{N} \qty(0.1, 0.1)
  \end{split}
\end{align}
where $\xi$ is an estimated coefficient of variation.
The stan models used are available with other codes at \url{https://github.com/jdossgollin/2018-robust-adaptation-cyclical-risk}.

@@

In this section we describe the parameters used to generate the specific sets of sequences that are analyzed in the results section.
To explore the temporal structure of risk, we generate $J=1000$ sequences of streamflow, for each of several combinations of $M$ and $N$.
Streamflow sequences are generated using the stationary generating functions, both with and without superimposing a secular trend.

@@

In the main text we sampled climate risk by considering dependence on a single climate index, as shown in equation 4.
Here we examine the sensitivity of our results to this assumption by considering a different parameterization for $\mu(t)$, specifically a two-state Markov chain model.
A Markov chain explicitly models transition between a fixed number of regimes, mimicking similar phenomena observed in nature.
The transition matrix is given by
\begin{equation}
  P = \mqty[\pi_1 & 1-\pi_1 \\ 1-\pi_2 & \pi_2].
\end{equation}
The transition matrix is first used to generate a sequence of states $S(t)$.
The value $\mu(t)$ depends only on $S(t)$ and on time itself:
\begin{equation}
  \mu(t) = \begin{cases}
    \mu_{1} + \gamma_1 \qty(t - t_0) & \qqtext{if} S(t) = 1 \\
    \mu_{2} + \gamma_2 \qty(t - t_0) & \qqtext{if} S(t) = 2
  \end{cases}
\end{equation}
For simplicity, we assume that the coefficient of variation is the same for both states and that $\pi_1=\pi_2$.
We further impose $\mu_{1} > \mu_{2}$ so that state 1 can be interpreted as the ``wet'' state and state 2 as the ``dry'' state.
Parameters for the three experiments are given in Table \ref{tab:markov-stationary}.
Results for these experiments are shown in figures \ref{fig:secular-only-markov-bias-variance}, \ref{fig:lfv-markov-bias-variance}, and \ref{fig:lfv-secular-markov-bias-variance}.
